---
title: "Full RAG Pipeline with LLM Response"
format: html
execute:
  echo: true
---

```{r setup}
library(tidyverse)
library(text2vec)
library(RcppHNSW)
library(httr)
library(jsonlite)
library(knitr)

set.seed(123)
```

# 1. Load Wikipedia Documents
This example assumes a dataset of Wikipedia articles with at least:

- `id` — document ID  
- `title` — article title  
- `text` — article text  

# @ALI Replace this with whatever schema you make for the dataset

```{r}
# Placeholder example
dataset <- tibble(
  id = 1:5,
  title = c(
    "R (programming language)",
    "Quarto (software)",
    "Vector database",
    "Word embedding",
    "Retrieval-Augmented Generation"
  ),
  text = c(
    "R is a programming language and environment for statistical computing.",
    "Quarto is an open-source scientific and technical publishing system.",
    "A vector database enables efficient similarity search across embeddings.",
    "Word embeddings represent words as vectors in continuous space.",
    "Retrieval-Augmented Generation is a technique that improves LLM accuracy by adding external context."
  )
)

dataset
```

# 2. Create Embeddings for Documents

```{r}
tokens <- dataset$text %>% tolower() %>% word_tokenizer()

it <- itoken(tokens, progressbar = FALSE)
vocab <- create_vocabulary(it)
vectorizer <- vocab_vectorizer(vocab)

tcm <- create_tcm(it, vectorizer, skip_grams_window = 5)

glove <- GlobalVectors$new(rank = 300, x_max = 10)
glove_vectors <- glove$fit_transform(tcm, n_iter = 20)

emb_matrix <- glove_vectors + t(glove$components)
```

# 3. Convert Each Document into a Vector

```{r}
get_doc_embedding <- function(text, emb_matrix) {
  words <- word_tokenizer(tolower(text))[[1]]
  valid_words <- intersect(words, rownames(emb_matrix))

  if (length(valid_words) == 0) {
    return(rep(0, ncol(emb_matrix)))
  }

  colMeans(emb_matrix[valid_words, , drop = FALSE])
}

dataset$embedding <- lapply(dataset$text, get_doc_embedding, emb_matrix)
emb_mat <- do.call(rbind, dataset$embedding)
```

# 4. Build HNSW Vector Index

```{r}
index <- HNSW$new(M = 16, ef = 200, distance = "cosine")
index$add_items(emb_mat)
```

# 5. RAG Retrieval Function

```{r}
rag_retrieve <- function(query, top_k = 3) {
  q_emb <- get_doc_embedding(query, emb_matrix)
  result <- index$knn(q_emb, k = top_k)
  ids <- result$item_id

  tibble(
    rank = 1:top_k,
    id = dataset$id[ids],
    title = dataset$title[ids],
    text = dataset$text[ids],
    score = result$distance
  )
}
```

# 6. LLM Response Step (Using Retrieved Context)

```{r}
call_llm <- function(prompt, context) {
  api_key <- Sys.getenv("OPENAI_API_KEY") # Not sure if we should use personal or if we can get one for our group

  body <- list(
    model = "gpt-4.1-mini",
    messages = list(
      list(
        role = "system",
        content = "You are a helpful assistant who answers questions using the provided context."
      ),
      list(
        role = "user",
        content = paste(
          "Context:
",
          context,
          "

Question:
",
          prompt,
          sep = ""
        )
      )
    )
  )

  res <- httr::POST(
    url = "https://api.openai.com/v1/chat/completions",
    add_headers(
      Authorization = paste("Bearer", api_key),
      "Content-Type" = "application/json"
    ),
    body = toJSON(body, auto_unbox = TRUE)
  )

  parsed <- content(res, "parsed", simplifyVector = TRUE)
  parsed$choices[[1]]$message$content
}
```

# 7. Complete RAG Pipeline Wrapper

```{r}
rag_pipeline <- function(query, top_k = 3) {
  retrieved <- rag_retrieve(query, top_k)

  combined_context <- paste(
    paste0(
      "### ",
      retrieved$title,
      "
",
      retrieved$text
    ),
    collapse = "

"
  )

  answer <- call_llm(query, combined_context)

  list(
    retrieved = retrieved,
    answer = answer
  )
}
```

# 8. Example Query

```{r}
result <- rag_pipeline("What is retrieval-augmented generation?")
result$retrieved %>% kable()
result$answer
```

# 9. Interactive Prompt

```{r}
query <- readline("Ask a question about Wikipedia content: ")
result <- rag_pipeline(query)

cat("\n\n--- Retrieved Documents ---\n")
print(kable(result$retrieved))

cat("\n\n--- LLM Answer ---\n")
cat(result$answer)
```
