---
title: "Ask Wikipedia (Powered by LLM)"
format: html
server: shiny
---


```{r}
install.packages("shiny", "httr", "openai")
```
```{r}
library(shiny)
library(httr)
library(openai)

# -------------------------------------------------------
# Load API Key
# -------------------------------------------------------
key_path <- "openai_key.txt"
if (file.exists(key_path)) {
  openai_key <- readLines(key_path, warn = FALSE)
  Sys.setenv(OPENAI_API_KEY = openai_key)
  print("Loaded OpenAI API key from file.")
} else {
  openai_key <- Sys.getenv("OPENAI_API_KEY")
  print("Loaded OpenAI API from system variable.")
}

if (openai_key == "") {
  stop("OpenAI API key not found. Please set it in 'openai_key.txt' or as an environment variable 'OPENAI_API_KEY'.")
}
# -------------------------------------------------------
# Rewrite query using LLM for better Wikipedia search
# -------------------------------------------------------
rewrite_query <- function(question) {
  response <- create_chat_completion(
    model = "gpt-4.1-mini",
    messages = list(
      list(
        role = "system",
        content = "Rewrite the user's question as a short, clean keyword-style query for Wikipedia. Return ONLY the query."
      ),
      list(role = "user", content = question)
    )
  )

  query <- response$choices[1, "message.content"]
  query <- gsub("[^a-zA-Z0-9 \\-]", "", query)
  query <- trimws(query)

  cat("=== DEBUG: rewritten query ===\n")
  print(query)

  return(query)
}

# -------------------------------------------------------
# Fetch raw wikitext from Wikipedia using MediaWiki API
# -------------------------------------------------------
fetch_wikitext <- function(title) {
  url <- "https://en.wikipedia.org/w/api.php"

  res <- httr::GET(
    url,
    query = list(
      action = "query",
      prop = "revisions",
      titles = title,
      rvprop = "content",
      rvslots = "main",
      format = "json"
    )
  )

  json <- httr::content(res, "parsed")
  page <- json$query$pages[[1]]

  text <- page$revisions[[1]]$slots$main$`*`

  return(text)
}

# -------------------------------------------------------
# Search Wikipedia + fetch wikitext for each result
# -------------------------------------------------------
fetch_wikipedia <- function(query, limit = 5) {
  cat("=== DEBUG: starting wikipedia search ===\n")

  url <- "https://en.wikipedia.org/w/api.php"
  res <- httr::GET(
    url,
    query = list(
      action = "query",
      list = "search",
      srsearch = query,
      srlimit = limit,
      format = "json"
    )
  )

  json <- httr::content(res, "parsed")

  cat("=== DEBUG: raw search response ===\n")
  print(json$query$search)

  titles <- sapply(json$query$search, function(x) x$title)

  cat("=== DEBUG: search result titles ===\n")
  print(titles)

  pages <- lapply(titles, function(title) {
    wt <- fetch_wikitext(title)
    if (is.null(wt)) {
      wt <- ""
    }

    list(
      title = title,
      url = paste0("https://en.wikipedia.org/wiki/", gsub(" ", "_", title)),
      text = wt
    )
  })

  cat("=== DEBUG: wikitext character lengths ===\n")
  print(sapply(pages, function(p) nchar(p$text)))

  return(pages)
}

# -------------------------------------------------------
# LLM Summarization
# -------------------------------------------------------
summarize_with_llm <- function(question, pages) {
  retrieved_text <- paste(
    sapply(pages, function(p) {
      paste0(
        "### ",
        p$title,
        "\n",
        substr(p$text, 1, 1200),
        "...\n"
      )
    }),
    collapse = "\n\n"
  )

  prompt <- paste0(
    "You are assisting a user by answering their question using ONLY the Wikipedia excerpts provided.\n\n",
    "User question: ",
    question,
    "\n\n",
    "Wikipedia excerpts:\n\n",
    retrieved_text,
    "\n\n",
    "Using ALL of the provided Wikipedia excerpts, write a single unified answer.\n",
    "Synthesize the information across all pages.\n",
    "Do NOT list separate excerpts.\n\n",
    "At the end, include a Citations section using ONLY the pages provided.\n",
    "Format each line as:\n",
    "- TITLE — URL\n"
  )

  cat("=== DEBUG: prompt sent to LLM ===\n")
  cat(prompt, "\n\n")

  response <- create_chat_completion(
    model = "gpt-4.1",
    messages = list(
      list(
        role = "system",
        content = "You accurately summarize Wikipedia content."
      ),
      list(role = "user", content = prompt)
    )
  )

  answer_raw <- response$choices[1, "message.content"]

  cat("=== DEBUG: raw LLM output ===\n")
  print(answer_raw)

  return(answer_raw)
}

# -------------------------------------------------------
# Extract citations
# -------------------------------------------------------
extract_citations <- function(answer_text) {
  lines <- unlist(strsplit(answer_text, "\n"))
  cite_lines <- grep("^\\- ", lines, value = TRUE)

  citations <- lapply(cite_lines, function(line) {
    parts <- unlist(strsplit(sub("^\\- ", "", line), " — "))
    list(
      title = parts[1],
      url = parts[2]
    )
  })

  return(citations)
}

# -------------------------------------------------------
# Remove citations section before printing
# -------------------------------------------------------
clean_answer <- function(answer_text) {
  lines <- unlist(strsplit(answer_text, "\n"))
  idx <- grep("^Citations", lines)

  if (length(idx) > 0) {
    cleaned <- lines[1:(idx[1] - 1)]
  } else {
    cleaned <- lines
  }

  paste(cleaned, collapse = "\n")
}

# -------------------------------------------------------
# UI
# -------------------------------------------------------
ui <- fluidPage(
  tags$style(HTML(
    "
    .answer-box pre {
      white-space: pre-wrap !important;
      word-break: break-word !important;
    }
  "
  )),

  titlePanel("Ask Wikipedia (Powered by LLM)"),

  textInput("question", "Enter a question:", width = "60%"),
  actionButton("ask", "Ask"),
  hr(),

  h3("Answer"),
  div(class = "answer-box", verbatimTextOutput("answer")),

  h3("Citations"),
  uiOutput("citations")
)

# -------------------------------------------------------
# SERVER
# -------------------------------------------------------
server <- function(input, output, session) {
  results <- eventReactive(input$ask, {
    req(input$question)

    query <- rewrite_query(input$question)
    pages <- fetch_wikipedia(query)

    answer_raw <- summarize_with_llm(input$question, pages)

    answer_clean <- clean_answer(answer_raw)
    citations <- extract_citations(answer_raw)

    list(
      answer = answer_clean,
      citations = citations
    )
  })

  output$answer <- renderText({
    req(results())
    results()$answer
  })

  output$citations <- renderUI({
    req(results())
    lapply(results()$citations, function(c) {
      tags$div(tags$a(href = c$url, c$title, target = "_blank"))
    })
  })
}

shinyApp(ui, server)
